---
title: "VideoGameSales"
author: "Cody B"
date: "`r Sys.Date()`"
output: github_document
---

This analysis examines trends in video game sales by analyzing data from titles 
released between 1971 and 2024. Specifically, it focuses on the total number of 
video game releases and the average sales figures across different genres. 
Additionally, the study explores sales data for video games by console, analyzing 
both the total releases and average sales per platform

The "Video Game Sales" data set can be found here: 
<https://mavenanalytics.io/data-playground?page=2&pageSize=5>.


```{r, echo=FALSE, include=FALSE}
library(tidyverse)
```

```{r, echo=FALSE}
vgchartz_2024 <- read_csv("VideoGameSales/vgchartz-2024.csv", col_types = cols(img = col_skip()))
```


## Genre


```{r echo=FALSE}
ggplot(data = vgchartz_2024) +
  geom_bar(mapping = aes(x = genre), stat = "count", fill = "cadetblue4", colour = "black") +
  scale_x_discrete(limits = names(sort(table(vgchartz_2024$genre), decreasing = TRUE))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Number Of Video Games Released By Genre", x = "Genre", y = "Number Of Video Games Released")
```

As demonstrated, the "Miscellaneous" genre has the highest total video game releases 
from 1971 to 2024. Following closely are the "Action," "Adventure," "Role Playing," 
and "Sports" genres.

To gain deeper insights, it would be valuable to further analyze the "Miscellaneous" 
genre to better understand the specific types of games included within it.


```{r echo=FALSE}
mean_sales <- vgchartz_2024 %>%
  group_by(genre) %>%
  summarize(mean_sales = mean(total_sales, na.rm = T), n_obs = n(), .groups = "drop") %>%
  arrange(desc(mean_sales))
  
ggplot(data = mean_sales) +
  geom_bar(mapping = aes(x = reorder(genre, mean_sales, decreasing = T), y = mean_sales), stat = "identity", fill = "cadetblue4", colour = "black") +
  geom_text(aes(x = reorder(genre, mean_sales, decreasing = TRUE), y = mean_sales, label = paste(n_obs)), vjust = -0.5, colour = "black", size = 3) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Mean Total Sales By Genre", x = "Genre", y = "Mean Total Sales (millions)") + 
  annotate("text", x = 7, y = .7, label = "Number of observations above bars", size = 3)
```

On average, video games categorized as "Sandbox" generate the highest total sales 
compared to other genres. The "Shooter," "Action-Adventure," "Sports," and "Action" 
genres also demonstrate strong sales performance.

## Console

```{r echo=FALSE}
filtered_data <- vgchartz_2024 %>%
  group_by(console) %>%
  summarize(n_obs = n(), .groups = "drop") %>%
  filter(n_obs > 750)
  
ggplot(data = filtered_data) +
  geom_bar(mapping = aes(x = reorder(console, n_obs, decreasing = T), y = n_obs), stat = "identity", fill = "cadetblue4", colour = "black") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(title = "Number of Video Games Released By Console (> 750)", x = "Console", y = "Number of Video Games Released")
```

Between 1971 and 2024, more video games were released for PC than for any other console. 
Notably, it also appears that PlayStation consoles have seen more game releases than Xbox consoles.

```{r, echo=FALSE}
mean_sales <- vgchartz_2024 %>%
  group_by(console) %>%
  summarize(mean_sales = mean(total_sales, na.rm = T), n_obs = n(), .groups = "drop") %>%
  filter(mean_sales > 0.18) %>%
  arrange(desc(mean_sales))
  
ggplot(data = mean_sales) +
  geom_bar(mapping = aes(x = reorder(console, mean_sales, decreasing = T), y = mean_sales), stat = "identity", fill = "cadetblue4", colour = "black") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(title = "Mean Total Sales By Console (> 0.18)", x = "Console", y = "Mean Total Sales (millions)")
```

On average, video games released for the Gameboy Color sold more copies than those 
on any other console.
